---
title: "Supplementary Materials"
author: "Stephanie Kobakian, Mitchell O'Hara-Wild, Dianne Cook, Stephanie Kovalcik"
date: "13 February 2017"
output: bookdown::pdf_document2
fontsize: 11pt
documentclass: article
bibliography: references.bib
---


```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

library(bookdown)
library(pander)
library(ggplot2)
library(imager)
library(devtools)
library(knitcitations)
library(RefManageR)
library(readr)
library(knitr)
library(grid)
library(tidyverse)
library(magick)
library(kfigr)
library(tables)
library(UpSetR)
library(descr)

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.env = TRUE)
options("citation_format" = "pandoc")

BibOptions(check.entries = FALSE, style = "markdown", bib.style = "alphabetic", cite.style = 'alphabetic')

#csv files:
ALLmetaIMG<-read_csv("ALLmetaIMG1.csv", col_types = cols(type = col_factor(levels = c("Manual", "Animetrics", "Google", "Microsoft", "Skybiometry"))))
SceneAttributes<-read_csv("TableCSVs/SceneAttributes.csv")
FaceAttributes<-read_csv("TableCSVs/FaceAttributes.csv")
SolutionSpecs<-read_csv("TableCSVs/SolutionSpecifications.csv")


#plot images with overlaid boxes
overlayGgplot <- function(imgList, mergeData, matchBox=FALSE){
  for(i in imgList){
    
    img <- image_read(paste0("images/", i))
    
    p <- ggplot(data.frame(x=c(0,800), y=c(0,450)), aes(x,y)) +
      theme_void()+ coord_fixed()+
      annotation_custom(rasterGrob(img)) +
      scale_x_continuous(expand = c(0,0), limits = c(0,800)) +
      scale_y_continuous(expand = c(0,0), limits = c(0,450))
    
    faceData <- ALLmetaIMGFaces %>% filter(file == i)
    
    if(nrow(faceData) == 0){
      return(p)
    } else{
      
      
      faceData <- faceData %>%
        mutate(x1 = minX, x2 = minX, x3 = maxX, x4 = maxX, x5 = minX,
               y1 = minY, y2 = maxY, y3 = maxY, y4 = minY, y5 = minY) %>%
        gather(corner, boxPos, x1:y5) %>%
        mutate(posCoord = substr(corner, 1, 1), posID = substr(corner, 2, 2)) %>%
        dplyr::select(-corner) %>% spread(posCoord, boxPos)
      
    }
    if(matchBox){
      p <- p + geom_path(aes(x=x, y=450-y, group = c(type, boxID), colour=boxID), faceData) + guides(colour="none")
    }
    else{
      p <- p + geom_path(aes(x=x, y=450-y, group = c(FaceID), colour=type), faceData) + scale_color_manual(values = c(Skybiometry="#ffe093", Microsoft="#5cc3f6", Google="#5cba82", Animetrics="#FF959c", Manual="#db70ff"))
    }
  return(p)
  }
}


# Create graphs of factors in manual proportions
getManualCount <- function(type, count) {
  return(count[type == "Manual"])
}
ggplotProportion <- function(dataset, factorVar){
  factorVar <- deparse(substitute(factorVar))
  dataset <- dataset %>% filter(matchesManual) %>% group_by_(factorVar, "type") %>% summarise(nTotal=n()) %>% group_by_(factorVar) %>% mutate(ManualCount = getManualCount(type, nTotal)) %>%
    mutate(proportion = nTotal/ManualCount) %>% rename_(xvar = factorVar) %>% filter(type!="Manual")
  ggplot(dataset, aes(x=factor(xvar), y=proportion, group = type, fill=type)) + geom_bar(stat = "identity", position = "dodge") +
    ylab("Proportion of faces matched") + xlab(factorVar)
}



#Factor Names 
ALLmetaIMG$graphic<-factor(ALLmetaIMG$graphic, levels = 0:1, labels = c("Live image", "Graphic"))
ALLmetaIMG$bg<-factor(ALLmetaIMG$bg, levels = 0:3, labels = c("Crowd", "Court", "Logo wall", "Not applicable"))
ALLmetaIMG$person<-factor(ALLmetaIMG$person, levels = 0:1, labels = c("No Person", "Person"))
ALLmetaIMG$shotangle<-factor(ALLmetaIMG$shotangle, levels = 0:2, labels = c("Player Shoulder Height", "Birds Eye", "Upward Angle"))
ALLmetaIMG$situation<-factor(ALLmetaIMG$situation, levels = 0:5, labels = c("Court in play", "Court player close-up", "Court close-up not player", "Crowd", "Off court close up of player", "Transition"))

ALLmetaIMG$detect<-factor(ALLmetaIMG$detect, levels = 0:3, labels = c("Player", "Other staff on court", "Fan", "None"))
ALLmetaIMG$obscured<-factor(ALLmetaIMG$obscured, levels = 0:1, labels = c("No", "Yes"))
ALLmetaIMG$lighting<-factor(ALLmetaIMG$lighting, levels = 0:2, labels = c("Direct sunlight", "Shaded", "Partially shaded"))
ALLmetaIMG$headangle<-factor(ALLmetaIMG$headangle, levels = 0:3, labels = c("Front on", "Back of head", "Profile", "Other"))
ALLmetaIMG$glasses<-factor(ALLmetaIMG$glasses, levels = 0:1, labels = c("No", "Yes"))
ALLmetaIMG$visorhat<-factor(ALLmetaIMG$visorhat, levels = 0:1, labels = c("No", "Yes"))

ALLmetaIMG$headangle <- as.character(ALLmetaIMG$headangle)
ALLmetaIMG$headangle[ALLmetaIMG$headangle == "Back of head"] <- "Other"
ALLmetaIMG$headangle <- as.factor(ALLmetaIMG$headangle)

ALLmetaIMGPlayers<- ALLmetaIMG %>% filter(detect=="Player")

# change this to facesVenn with binary information
ALLmetaIMGFaces<-ALLmetaIMG%>%
  filter(matchesManual) %>% 
  dplyr::select(-ID, -facecounter)%>%
  mutate(fileID = as.numeric(factor(file))) %>%
  mutate(FaceKey=paste(fileID, boxID, sep="-")) %>%
  mutate(FaceID=paste(fileID, boxID, substring(type, 1,2), sep="-"))


```

Talk about how sample was created?


The Shiny [@shinyw] package provides a 'framework for building web applications using R'. It allows for the creation of web applications using R rather than 'requiring HTML, CSS, or JavaScript knowledge.'


The program we created was an interactive application that worked by presenting an image, using the imager @imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attributes in Scene-Attributes-Table.


Describe a hit in regression 

```{r Regression, results='hide', fig.cap = "\\label{Regression}"}
library(purrr)

hitmiss <- function(x){
  allType <- c("Animetrics", "Google", "Microsoft", "Skybiometry")
  hit <- allType %in% x$type
  x[1,] %>%
    dplyr::select(file:visorhat) %>%
    cbind(type = allType, hit = hit)
}

glmFits <- ALLmetaIMGFaces %>%
  split(.$FaceKey) %>%
  map_df(~ hitmiss(.)) %>%
  split(.$type) %>%
  map(~ glm(hit ~ 0+., data = select(., -type, -file)))

glmSummary <- glmFits %>%
  map(~ rename(cbind(rownames_to_column(cbind(as.data.frame(coef(summary(.)))))), variable = rowname))

glmPlot <- do.call(rbind, Map(cbind, glmSummary, type = names(glmSummary)))

# Graphs
## Coefficient of variables by software
#glmPlot %>%
#  ggplot(aes(x=type, y=Estimate)) +
#  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
#  geom_col() + 
#  facet_wrap(~ variable)

## Significance of variables by software
glmPlot %>%
  mutate(significant = `Pr(>|t|)` < 0.05) %>%
  mutate(`Pr(<|t|)` = 1 - `Pr(>|t|)`) %>%
  ggplot(aes(x=type, y=`Pr(<|t|)`)) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  geom_col(aes(fill=significant)) + 
  facet_wrap(~ variable)

## Significance of Estimated variable influences by software
glmPlot %>%
  mutate(significant = `Pr(>|t|)` < 0.05) %>%
  ggplot(aes(x=type, y=Estimate)) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  geom_col(aes(fill=significant)) + 
  facet_wrap(~ variable)


```


```{r Regression-Table, fig.cap = "\\label{Regression-Table}"}

RegSituation<-glmPlot %>% subset(grepl("situation*", variable)) %>% select(variable, Estimate, type)

#RegSituation


data_wide <- spread(RegSituation, variable, Estimate)

names(data_wide)[names(data_wide)=="situationCourt player close-up"] <- "Court player close-up"
names(data_wide)[names(data_wide)=="situationCrowd"] <- "Crowd"
names(data_wide)[names(data_wide)=="situationOff court close up of player"] <- "Off court close up of player"
names(data_wide)[names(data_wide)=="situationCourt close-up not player"] <- "Court close-up not player"

kable(data_wide, format = "markdown", booktabs = TRUE, caption = "")
```



```{r Regression-Table, fig.cap = "\\label{Regression-Table}"}

RegHeadAngle<-glmPlot %>% subset(grepl("headangle*", variable)) %>% select(variable, Estimate, type)

#RegSituation


data_wideHead <- spread(RegHeadAngle, variable, Estimate)

names(data_wideHead)[names(data_wideHead)=="headangleOther"] <- "Other"
names(data_wideHead)[names(data_wideHead)=="headangleProfile"] <- "Profile"

kable(data_wideHead, format = "markdown", booktabs = TRUE, caption = "")
```





```{r Caret, fig.cap = "\\label{Caret}"}
library(caret)
library(e1071)
set.seed(100)

ALLmetaIMGFaces
inTrain <- createDataPartition(y = ALLmetaIMG$matchesManual, p = .75, 
                               list = FALSE)

training <- ALLmetaIMG[ inTrain,]
testing <- ALLmetaIMG[-inTrain,]



ctrl <- trainControl(method = "repeatedcv",
repeats = 3,
classProbs = TRUE,
summaryFunction = twoClassSummary)

plsFit <- train(hit ~ .,
 data = training,
 method = "pls",
 ## Center and scale the predictors for the training
 ## set and all future samples.
 preProc = c("center", "scale"), metric="Accuracy")

# plsFit <- train(hit ~ .,
# data = training,
# method = "glm", 
# tuneLength = 15,
# trControl = ctrl,
# metric = "ROC",
# preProc = c("center", "scale"))
plsFit

```

